---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv('activity.csv')
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activity$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activity$interval), format='%H:%M')
```

-----

## What is mean total number of steps taken per day?
```{r}
StepPerDay <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(StepPerDay, xlab='Steps per day', ylab='Frequency - binwidth 500', binwidth=500)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
PerDayMean <- mean(StepPerDay)
PerDayMedian <- median(StepPerDay)
```
* Mean: `r PerDayMean`
* Median:  `r PerDayMedian`

-----

## What is the average daily activity pattern?
```{r}
AverageStepsDaily <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r}
ggplot(data=AverageStepsDaily, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("Interval - 5min") +
    ylab("Average number of steps") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
MaxSteps <- which.max(AverageStepsDaily$meanSteps)
MostTime <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", AverageStepsDaily[MaxSteps,'interval'])
```

* Most Steps at: `r MostTime`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
MissingValuesCount <- length(which(is.na(activity$steps)))
```

* Number of missing values: `r MissingValuesCount`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.

* Replace missing values with the mean steps 

##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityImpute <- activity
activityImpute$steps <- impute(activity$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
StepsPerDayImpute <- tapply(activityImpute$steps, activityImpute$date, sum)
qplot(StepsPerDayImpute, xlab='Steps per day (Imputed)', ylab='Frequency - binwidth 500', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
StepsPerDayMeanImpute <- mean(StepsPerDayImpute)
StepsPerDayMedianImpute <- median(StepsPerDayImpute)
```
* Mean (Imputed): `r StepsPerDayMeanImpute`
* Median (Imputed):  `r StepsPerDayMedianImpute`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activityImpute$dateType <-  ifelse(as.POSIXlt(activityImpute$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
AvgActivityImpute <- aggregate(steps ~ interval + dateType, data=activityImpute, mean)
ggplot(AvgActivityImpute, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
